Automatically Splitting a Two-Stage Lambda Calculus

### 2015-07-11 paper/conclusion.tex
reviewer comments
compressed figures

This paper presents techniques to translate staged lambda calculus
programs to multi-pass programs, which evaluate in a sequence of
passes, each of which is a conventional program.  We present a
formalization of the techniques in a way that can be implemented in
practice.  While we have also provided an implementation that
successfully splits the examples considered, we have stopped short of
offering a full proof of correctness, which we leave for future work.
We are also interested in providing a more complete implementation and
investigating how these techniques can be integrated into existing
practical systems such as existing staged programming systems for
computer graphics.  On the more conceptual side, we plan to extend the
techniques to richer source languages, for example those that include
parallelism.

>>>>

This paper presents a splitting algorithm for \lang, 
a two-staged typed lambda calculus with support for recursion and first-class functions.
The type system of \lang\ uses three worlds \bbonem, \bbonep, and \bbtwo\ to classify code by its stage,
with the modalities $\fut$ and $\curr$ providing internalizations 
of second-stage code and ground first-stage code to the mixed world.
We present a dynamic semantics that evaluates terms in two passes
and provides an eager interpretation of {\tt next} via hoisting steps.
We prove the correctness of our splitting algorithm against the dynamic semantics,
and implement this proof in Twelf.
Finally, we discuss our implementation of \lang\ and its splitting algorithm
and analyze their behavior for a number of staged programs,
including those with higher-order and recursive functions.

Looking forward, we are interested in investigating the behavior of splitting 
in the presence of richer language features such as mutation, polymorphism,
parallel constructs, and more than two stages.
Also while we have proven the dynamic correctness of our splitting algorithm,
we do not yet have a characterization of the types of the output
or of the cost behavior of output terms.
